---
title: 'TF-IDF & Sentiment Analysis '
author: Minr
date: '2020-06-09'
slug: sentiment-analysis
categories: []
tags:
  - NLP
subtitle: ''
summary: ''
authors: []
lastmod: '2020-06-09T22:46:01-07:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


The TF-IDF is computed by multiplying two terms:TF and IDF. TF is the normalized term frequency and IDF is the inverse document frequency.
Therefore, the formula for TF-IDF is TFxIDF 

where,
$$\text{TF} = (\frac{\text{number of times term appears in a document}}{\text{Total number of terms in the document}})$$
and
$$\text{IDF} = \ln(\frac{n_{\text{ documents }}}{n_{\text{document containing term}}})$$
TF-IDF can be used for ranking important documents(page ranking). For example when we google a term COVID-19, the results are displayed in order of relevance. TF-IDF gives searched word(COVID-19) a higher score. 

### TF-IDF on Gutenberg books

```{r}
library(pacman)
p_load(gutenbergr,tidyr,tidyverse,tidytext,forcats,
       genius)
```

```{r}
#books on mathematics from project glutenberg
math_books <- gutenberg_download(c(16449,25664,17001,16713), meta_fields = "title")
```


```{r}
#frequencies of words in each book
math_words <- math_books %>% 
  unnest_tokens(word, text) %>% count(title, word, sort = TRUE)
math_words
```


```{r}
#tf-idf
tfidf_mathbooks <- math_words %>%
  bind_tf_idf(word, title, n) %>%
  mutate(word = fct_reorder(word, tf_idf)) %>%
  mutate(title = factor(title, levels = c( "The Number Concept: Its Origin and Development","The Earliest Arithmetics in English", "An Elementary Course in Synthetic Projective Geometry","Amusements in Mathematics")))
tfidf_mathbooks
```



```{r}
#bargraphs
tfidf_mathbooks %>% 
  group_by(title) %>% 
  top_n(15, tf_idf) %>% 
  ungroup() %>%
  mutate(word = reorder(word, tf_idf)) %>%
  ggplot(aes(word, tf_idf, fill = title)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~title, ncol = 2, scales = "free") +
  coord_flip() + ggtitle("Bargraphs of top 15 words")
```


Amusements in Mathematics - This book appears to be related to solving puzzle and game.

An Elementary Course in Synthetic Projective Geometry- It appears to be related to book about measuring angle and point, probably involve a lot of theories.

The Earliest Arithmetics in English - The terms are interesting. The words in various langauge in original form. 

The Number Concept: Its Origin and Development - It appears to be related how counting and numbering originated. It is probably indicating use of fingers by tribes for counting.



***Same analysis using bigrams.**


```{r}
# bigrams for the math books from previous problem
math_bigrams <-  math_books %>%
unnest_tokens(bigram, text, token = "ngrams", n = 2)
math_bigrams
```

```{r}
# count bigrams
math_bigrams %>% count(bigram, sort = TRUE)
```


```{r}
# separate and filter stop words
mathbigrams_separated <- math_bigrams %>%
separate(bigram, c("word1", "word2"), sep = " ")

mathbigrams_filtered <- mathbigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

# new bigram counts
bigram_counts <- mathbigrams_filtered %>% 
  count(word1, word2, sort = TRUE)
bigram_counts

```


```{r}
#combine filtered words to form bigrams 
mathbigrams_united <- mathbigrams_filtered %>%
  unite(bigram, word1, word2, sep = " ")
mathbigrams_united
```

```{r}
bigram_tf_idf <- mathbigrams_united %>%
  count(title, bigram) %>%
  bind_tf_idf(bigram, title, n) %>%
  arrange(desc(tf_idf))
bigram_tf_idf
```


```{r}
#bargraphs
bigram_tf_idf %>% 
  group_by(title) %>% 
  top_n(15, tf_idf) %>% 
  ungroup() %>%
  mutate(bigram=reorder(bigram, tf_idf)) %>%
  ggplot(aes(bigram, tf_idf, fill = title)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~title, ncol = 2, scales = "free") +
  coord_flip() + ggtitle("Bargraphs of top 15 bigrams")
```



### Sentiment analysis on the songs from Genius

1. Aerosmith - I don't want to miss a thing

```{r}
#get song lyrics
aero_smith_i_don <- genius_lyrics(artist = "Aerosmith", song = " I don't want to miss a thing")
aero_smith_i_don
```


```{r}
#tidy up lyrics
i_dont_tidy <- aero_smith_i_don  %>% select(lyric, track_title) %>% unnest_tokens(word, lyric)
i_dont_tidy
```

```{r}
# join with sentiment lexicon
i_dont_sentiments<- i_dont_tidy%>%
inner_join(get_sentiments("bing"), by = c(word = "word"))
i_dont_sentiments
```

```{r}
#bargraph word-sentiment
i_dont_sentiments %>%
  count(sentiment, word) %>%
  ungroup() %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  ylab("Contribution to sentiment") +
  coord_flip()

```


2. The Weeknd - Earned It

```{r}
the_weeknd_earned_it <- genius_lyrics(artist = "The Weeknd", song = "Earned It")
the_weeknd_earned_it
```


```{r}
earned_it_tidy <- the_weeknd_earned_it  %>% select(lyric, track_title) %>% unnest_tokens(word, lyric)
earned_it_tidy
```

```{r}
earned_it_sentiments<- earned_it_tidy%>%
inner_join(get_sentiments("bing"), by = c(word = "word"))
earned_it_sentiments
```

```{r}
earned_it_sentiments %>%
  count(sentiment, word) %>%
  ungroup() %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  ylab("Contribution to sentiment") +
  coord_flip()

```


3. Pharrell Williams - Happy

```{r}
pharrell_will_happy <- genius_lyrics(artist = "Pharrell Williams", song = "Happy")
pharrell_will_happy
```


```{r}
happy_tidy <- pharrell_will_happy  %>% select(lyric, track_title) %>% unnest_tokens(word, lyric)
happy_tidy
```

```{r}
happy_sentiments<- happy_tidy%>%
inner_join(get_sentiments("bing"), by = c(word = "word"))
happy_sentiments
```

```{r}
happy_sentiments %>%
  count(sentiment, word) %>%
  ungroup() %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  ylab("Contribution to sentiment") +
  coord_flip()

```



4.  Summertime Sadness – Lana Del Rey

```{r}
 Summertime_Sadness <- genius_lyrics(artist = "Lana Del Rey", song = "Summertime Sadness")
Summertime_Sadness
```


```{r}
Summertime_Sadness_tidy <- Summertime_Sadness %>% select(lyric, track_title) %>% unnest_tokens(word, lyric)
Summertime_Sadness_tidy
```

```{r}
Summertime_Sadness_sentiments<- Summertime_Sadness_tidy%>%
inner_join(get_sentiments("bing"), by = c(word = "word"))
Summertime_Sadness_sentiments
```

```{r}
Summertime_Sadness_sentiments %>%
  count(sentiment, word) %>%
  ungroup() %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  ylab("Contribution to sentiment") +
  coord_flip()

```




