---
title: 'Sentiment Analysis with songs '
author: Minr
date: '2020-06-11'
slug: sentiment-analysis
categories: []
tags:
  - NLP
subtitle: ''
summary: ''
authors: []
lastmod: '2020-06-11T22:46:01-07:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


**Sentiment analysis** is the interpretation and classification of emotions (positive, negative and neutral) within text data using natural language processing(NLP) techniques. 

In this session, we are going to use songs from R package genius with bing lexicon. The bing lexicon categorizes words in a binary fashion into positive and negative categories with weighted sentiment scores.

First, load libraries
```{r}
library(pacman)
p_load(tidyr,tidyverse,tidytext,forcats, genius)
```

then download songs using genius functions and perform sentiment analysis. Steps are described below.

1. Aerosmith - I don't want to miss a thing

```{r}
#get song lyrics
aero_smith_i_don <- genius_lyrics(artist = "Aerosmith", song = " I don't want to miss a thing")
aero_smith_i_don
```


```{r}
#tidy up lyrics
i_dont_tidy <- aero_smith_i_don  %>% select(lyric, track_title) %>% unnest_tokens(word, lyric)
i_dont_tidy
```

```{r}
# join with sentiment lexicon
i_dont_sentiments<- i_dont_tidy%>%
inner_join(get_sentiments("bing"), by = c(word = "word"))
i_dont_sentiments
```

```{r}
#bargraph word-sentiment
i_dont_sentiments %>%
  count(sentiment, word) %>%
  ungroup() %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  ylab("Contribution to sentiment") +
  coord_flip()

```


2. The Weeknd - Earned It

```{r}
the_weeknd_earned_it <- genius_lyrics(artist = "The Weeknd", song = "Earned It")
the_weeknd_earned_it
```


```{r}
earned_it_tidy <- the_weeknd_earned_it  %>% select(lyric, track_title) %>% unnest_tokens(word, lyric)
earned_it_tidy
```

```{r}
earned_it_sentiments<- earned_it_tidy%>%
inner_join(get_sentiments("bing"), by = c(word = "word"))
earned_it_sentiments
```

```{r}
earned_it_sentiments %>%
  count(sentiment, word) %>%
  ungroup() %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  ylab("Contribution to sentiment") +
  coord_flip()

```


3. Pharrell Williams - Happy

```{r}
pharrell_will_happy <- genius_lyrics(artist = "Pharrell Williams", song = "Happy")
pharrell_will_happy
```


```{r}
happy_tidy <- pharrell_will_happy  %>% select(lyric, track_title) %>% unnest_tokens(word, lyric)
happy_tidy
```

```{r}
happy_sentiments<- happy_tidy%>%
inner_join(get_sentiments("bing"), by = c(word = "word"))
happy_sentiments
```

```{r}
happy_sentiments %>%
  count(sentiment, word) %>%
  ungroup() %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  ylab("Contribution to sentiment") +
  coord_flip()

```



4.  Summertime Sadness – Lana Del Rey

```{r}
 Summertime_Sadness <- genius_lyrics(artist = "Lana Del Rey", song = "Summertime Sadness")
Summertime_Sadness
```


```{r}
Summertime_Sadness_tidy <- Summertime_Sadness %>% select(lyric, track_title) %>% unnest_tokens(word, lyric)
Summertime_Sadness_tidy
```

```{r}
Summertime_Sadness_sentiments<- Summertime_Sadness_tidy%>%
inner_join(get_sentiments("bing"), by = c(word = "word"))
Summertime_Sadness_sentiments
```

```{r}
Summertime_Sadness_sentiments %>%
  count(sentiment, word) %>%
  ungroup() %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  ylab("Contribution to sentiment") +
  coord_flip()

```


This is a glimpse of sentiment analysis and it's use. With few steps, one can analyse texts with positive and negative emotions in a manner general public can interpret. Imagine significance of sentiment analysis in social media monitoring, brand monitoring, market research and customer service.


